### 2016-12-23 TOMS_Paper/reviewers_answers.txt
Add sketch answers to referees

Referee: 1

Comments to the Author This paper is overall well written. Examples are very
illustrative and help understand the article. My only concern is about how
authors relates their theortical results with numerical performance examples.
The main focus of this paper is about transformations that reduces FLOPs.
Authors present elimination and reduction pattern of code and show that there is
an optimal loop transformation that reduces FLOPs. However, the numeric
performance results is about "speed-up" not about FLOPs. Also, the speed-up is
more significant when the polynomial degree is higher. This confuses me either
vectorization improves the performance or actual FLOPs decreases. My suggestions
are as follows: 1. It would be better to include actual transformed code and
original code in the appendix for all computing model (mass, helmholtz,
elasticity, hyper elasticity). At least I wonder what makes hugh speed up in the
hyper elastic problem compared to usual mass matrix.  2. Please include actual
FLOP measurements of input code and transformed code and show some correlation
with performance.

Referee: 2

Comments to the Author

Report on “An algorithm for the optimization of finite element integration
loops”

The paper is well written, interesting does clearly merit publication.  The
usage of advanced compiler technology as demonstrated here is clearly novel and
it brings accelerated performance.

There is one issue that is not covered in the analysis and only briefly
mentioned in the general discussion: the transfer from memory. I understand that
it complicates the analysis but the least it should be quantified in the
numerical experiments. As was shown by Alnæs and Mardal in 2010, TOMS, the
transfer is significant. This part of the story should also be told as I guess
the most common user, low-order elements with simple equation, will not notice
any benefit. This has been an untold story in several FEniCS papers. It is to a
certain extent discussed already in the paper but it should also be quantified.

Referee: 3

Comments to the Author The FEM is widely used in scientific computing and the
integration to compute the elements (local element matrices)  can often be time
consuming. This manuscript is about optimizing the finite element integration
loops and the number of flops therein. The authors propose an approach that is
shown to be locally optimal. An implementation (in COFFEE) is compared against
other methods and speedups over the default generated code is demonstrated in
several cases.

Overall this is a well written and interesting paper, even to non-experts.
However, as a non-expert, I have several concerns that I think should be
addressed:

1) The focus is on minimizing flops, but on modern architectures, memory
accesses are typically more expensive and flops are "almost free". Although
clearly of theoretical interest, it would be nice to justify that this
optimization also is important to reduce run time.

2) All experiments and results are run in serial on single core. In practice,
most computers today are multicore (or manycore) so some discussion about
extensions to multithreaded case would be appropriate. I recognize a parallel
implementation is beyond the scope of the current paper.

3) Results: The authors present several charts with speedups on different
problems, which gives detailed information. However, it would be nice to add a
figure with aggregate results, such as a scatter plot of all instances or a
performance profile.

4) The manuscript is already quite long so if possible, do not increase the
total length (i.e. try find things to tighten or cut).Referee: 1

Comments to the Author This paper is overall well written. Examples are very
illustrative and help understand the article. My only concern is about how
authors relates their theortical results with numerical performance examples.
The main focus of this paper is about transformations that reduces FLOPs.
Authors present elimination and reduction pattern of code and show that there is
an optimal loop transformation that reduces FLOPs. However, the numeric
performance results is about "speed-up" not about FLOPs. Also, the speed-up is
more significant when the polynomial degree is higher. This confuses me either
vectorization improves the performance or actual FLOPs decreases. My suggestions
are as follows: 1. It would be better to include actual transformed code and
original code in the appendix for all computing model (mass, helmholtz,
elasticity, hyper elasticity). At least I wonder what makes hugh speed up in the
hyper elastic problem compared to usual mass matrix.  2. Please include actual
FLOP measurements of input code and transformed code and show some correlation
with performance.

Referee: 2

Comments to the Author

Report on “An algorithm for the optimization of finite element integration
loops”

The paper is well written, interesting does clearly merit publication.  The
usage of advanced compiler technology as demonstrated here is clearly novel and
it brings accelerated performance.

There is one issue that is not covered in the analysis and only briefly
mentioned in the general discussion: the transfer from memory. I understand that
it complicates the analysis but the least it should be quantified in the
numerical experiments. As was shown by Alnæs and Mardal in 2010, TOMS, the
transfer is significant. This part of the story should also be told as I guess
the most common user, low-order elements with simple equation, will not notice
any benefit. This has been an untold story in several FEniCS papers. It is to a
certain extent discussed already in the paper but it should also be quantified.

Referee: 3

Comments to the Author The FEM is widely used in scientific computing and the
integration to compute the elements (local element matrices)  can often be time
consuming. This manuscript is about optimizing the finite element integration
loops and the number of flops therein. The authors propose an approach that is
shown to be locally optimal. An implementation (in COFFEE) is compared against
other methods and speedups over the default generated code is demonstrated in
several cases.

Overall this is a well written and interesting paper, even to non-experts.
However, as a non-expert, I have several concerns that I think should be
addressed:

1) The focus is on minimizing flops, but on modern architectures, memory
accesses are typically more expensive and flops are "almost free". Although
clearly of theoretical interest, it would be nice to justify that this
optimization also is important to reduce run time.

2) All experiments and results are run in serial on single core. In practice,
most computers today are multicore (or manycore) so some discussion about
extensions to multithreaded case would be appropriate. I recognize a parallel
implementation is beyond the scope of the current paper.

3) Results: The authors present several charts with speedups on different
problems, which gives detailed information. However, it would be nice to add a
figure with aggregate results, such as a scatter plot of all instances or a
performance profile.

4) The manuscript is already quite long so if possible, do not increase the
total length (i.e. try find things to tighten or cut).Referee: 1

Comments to the Author This paper is overall well written. Examples are very
illustrative and help understand the article. My only concern is about how
authors relates their theortical results with numerical performance examples.
The main focus of this paper is about transformations that reduces FLOPs.
Authors present elimination and reduction pattern of code and show that there is
an optimal loop transformation that reduces FLOPs. However, the numeric
performance results is about "speed-up" not about FLOPs. Also, the speed-up is
more significant when the polynomial degree is higher. This confuses me either
vectorization improves the performance or actual FLOPs decreases. My suggestions
are as follows: 1. It would be better to include actual transformed code and
original code in the appendix for all computing model (mass, helmholtz,
elasticity, hyper elasticity). At least I wonder what makes hugh speed up in the
hyper elastic problem compared to usual mass matrix.  2. Please include actual
FLOP measurements of input code and transformed code and show some correlation
with performance.

Referee: 2

Comments to the Author

Report on “An algorithm for the optimization of finite element integration
loops”

The paper is well written, interesting does clearly merit publication.  The
usage of advanced compiler technology as demonstrated here is clearly novel and
it brings accelerated performance.

There is one issue that is not covered in the analysis and only briefly
mentioned in the general discussion: the transfer from memory. I understand that
it complicates the analysis but the least it should be quantified in the
numerical experiments. As was shown by Alnæs and Mardal in 2010, TOMS, the
transfer is significant. This part of the story should also be told as I guess
the most common user, low-order elements with simple equation, will not notice
any benefit. This has been an untold story in several FEniCS papers. It is to a
certain extent discussed already in the paper but it should also be quantified.

Referee: 3

Comments to the Author The FEM is widely used in scientific computing and the
integration to compute the elements (local element matrices)  can often be time
consuming. This manuscript is about optimizing the finite element integration
loops and the number of flops therein. The authors propose an approach that is
shown to be locally optimal. An implementation (in COFFEE) is compared against
other methods and speedups over the default generated code is demonstrated in
several cases.

Overall this is a well written and interesting paper, even to non-experts.
However, as a non-expert, I have several concerns that I think should be
addressed:

1) The focus is on minimizing flops, but on modern architectures, memory
accesses are typically more expensive and flops are "almost free". Although
clearly of theoretical interest, it would be nice to justify that this
optimization also is important to reduce run time.

2) All experiments and results are run in serial on single core. In practice,
most computers today are multicore (or manycore) so some discussion about
extensions to multithreaded case would be appropriate. I recognize a parallel
implementation is beyond the scope of the current paper.

3) Results: The authors present several charts with speedups on different
problems, which gives detailed information. However, it would be nice to add a
figure with aggregate results, such as a scatter plot of all instances or a
performance profile.

4) The manuscript is already quite long so if possible, do not increase the
total length (i.e. try find things to tighten or cut).